Jawbone Up Wearer Performance Prediction
========================================================
-- Course Project for Practical Machine Learning



Project Description:

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

This human activity recognition research has traditionally focused on discriminating between different activities, i.e. to predict "which" activity was performed at a specific point in time (like with the Daily Living Activities dataset above). The approach we propose for the Weight Lifting Exercises dataset is to investigate "how (well)" an activity was performed by the wearer. The "how (well)" investigation has only received little attention so far, even though it potentially provides useful information for a large variety of applications,such as sports training.

In this work we first define quality of execution and investigate three aspects that pertain to qualitative activity recognition: the problem of specifying correct execution, the automatic and robust detection of execution mistakes, and how to provide feedback on the quality of execution to the user. We tried out an on-body sensing approach, but also an "ambient sensing approach" (by using Microsoft Kinect).

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

========================================================
Project Approaches:

First is to import the datasets, and set the dependent variable "classe" aside. 

```{r}
library(caret)
pml.training <- read.csv("C:/Users/BenBen/Documents/Google Drive/coursera-Data Science Specialization/08_PracticalMachineLearning/project/pml-training.csv", na.strings = c("NA", ""))
pml.testing <- read.csv("C:/Users/BenBen/Documents/Google Drive/coursera-Data Science Specialization/08_PracticalMachineLearning/project/pml-testing.csv", na.strings = c("NA", ""))
classe=pml.training$classe
```

The basic preprocessing

```{r}
### basic
#1 remove irrelevant
rIndex <- grep("X|user_name|cvtd_timestamp", names(pml.training))
pml.training <- pml.training[, -rIndex]
#2 remove zero var
nzv <- nearZeroVar(pml.training)
pml.training <- pml.training[, -nzv]
#3 remove na
NAs <- apply(pml.training, 2, function(x) {
  sum(is.na(x))
})
pml.training <- pml.training[, which(NAs == 0)]
pml.training0=pml.training
#4 factor to dummy **don't need
```

The advanced preprocessing (which is not used as they don't improve the prediction accuracy)
```{r}
### extra
#1 impute
preObj <- preProcess(pml.training[,-56],method="medianImpute")
pml.training<- predict(preObj,pml.training[,-56])
pml.training1=cbind(pml.training,classe) #best
#2 standardize
stdize=preProcess(pml.training[,-56],method=c("center","scale"))
pml.training=predict(stdize,pml.training[,-56])
pml.training2=cbind(pml.training,classe)
## check corr
M <- abs(cor(pml.training))
diag(M) <- 0
which(M > 0.9,arr.ind=T)
#3 pca
pca<- preProcess(pml.training,method="pca",thresh=0.9)
pml.training_=predict(pca,pml.training)
pml.training3=cbind(pml.training_,classe)
```


Here is the main model part. I used the two best performing models RF and GBM.
```
### model
set.seed(12345)
inTrain = createDataPartition(pml.training0$classe,p=3/4,list = FALSE)
pml.train<- pml.training0[inTrain, ]  
pml.test <- pml.training0[-inTrain,]  
modFit1<-train(pml.train$classe~.,data=pml.train,method="rf")
modFit2<-train(pml.train$classe~.,data=pml.train,method="gbm")
```


check the in-sample accuracy and out-of-sample accuracy.

```
#in sample accuracy
modFit1
modFit1$results
max(modFit1$results$Accuracy)
#out sample accuracy
pred <-predict(modFit1,pml.test)
pml.test$predRight<-pred==pml.test$classe
table(pred,pml.test$classe)
postResample(pred,pml.test$classe) #same1
confusionMatrix(pred,pml.test$classe) #same2
```


As the accuracy is very high, this model could apply to predict the 20 rows of new data. 
```
### pred
pred2<-predict(modFit1,pml.testing)
pml.testing.2<-pml.testing
pml.testing.2$classe<-pred2
```


Finally output the results.
```
### export
pml_write_files = function(x) {
  n = length(x)
  for (i in 1:n) {
    filename = paste0("C:/Users/BenBen/Documents/Google Drive/coursera-Data Science Specialization/08_PracticalMachineLearning/project/problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
answers <- pml.testing.2$classe
pml_write_files(answers)
```
